%* installation (deps, build/install procedure, making module/ml work)
%* configuration (either via ./configure at build time, or via 
%   $LMOD_FOO), most important configure options
%* setting up the system spider cache: configuring Lmod, updating spider 
%  cache, difference between system/user spider cache
%* + all (other) aspects mentioned on slide #26 in this slide deck: 
%   explained + example :)
%* what settarg is all about

\documentclass{beamer}

\usetheme[headernav]{TACC} %%Drop the 'headernav' if you don't like
                           %%the stuff at the top of your slide

\usepackage{amsmath,amssymb,amsthm}
\usepackage{alltt}
\usepackage{graphicx}

\title{Lmod}


\author{Robert McLay}
\institute{The Texas Advanced Computing Center}

\date{Jan. 30, 2015}  %% Use this if you want to fix the date in
                      %% stone rather than use \today

\newcommand{\bfnabla}{\mbox{\boldmath$\nabla$}}
\newcommand{\laplacian}[1]{\bfnabla^2 #1}
\newcommand{\grad}[1]{\bfnabla #1}
\newcommand{\tgrad}[1]{\bfnabla^T #1}
\newcommand{\dvg}[1]{\bfnabla \cdot #1}
\newcommand{\curl}[1]{\bfnabla \times #1}
\newcommand{\lap}[1]{\bfnabla^2 #1}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \begin{itemize}
    \item What are Environment Modules?
    \item What is Lmod?
    \item 
  \end{itemize}
\end{frame}



\section{Environment Modules}

\begin{frame}{What are Environment Modules?}
  \begin{itemize}
    \item A tool to set (and unset!) environment variables.
    \item Useful for adding elements to \$PATH, \$LD\_LIBRARY\_PATH
    \item Also remove elements from \$PATH, etc as well.
    \item 1st paper on Env. Modules in 1991 by Furlani.
    \item It is currently a TCL/C based tool (a.k.a Tmod)
  \end{itemize}
\end{frame}

\begin{frame}{Why is this useful?}
  \begin{itemize}
    \item High Performance Computers have hundreds of users.
    \item Physicists, Chemists, Biologist, Engineers need different
      software
    \item Modules provide a convenient way to support them all.
    \item A software developers delight: 
      \begin{itemize}
        \item Switch compilers easily.
        \item For both versions and programs.
      \end{itemize}
    \item 
  \end{itemize}
\end{frame}

\begin{frame}[fragile]
    \frametitle{An Example}
    {\small
      \begin{alltt}
          \$ {\color{blue} ddt}
          command not found: ddt
          \$ {\color{blue} module load ddt}
          \$ {\color{blue} ddt }
          \$ {\color{blue} module rm ddt}
          \$ {\color{blue} ddt}
          command not found: ddt
      \end{alltt}
    }
\end{frame}    

\begin{frame}{What is Lmod?}
  \begin{itemize}
    \item A Lua based Enviroment Module Tool that supports TCL modulefiles
    \item Tmod doesn't support a software hierarchy
    \item The C++ Boost library needs to match Compiler and Version
    \item Switch compilers should swap boost to match
    \item Lmod does this automatically
  \end{itemize}
\end{frame}

\begin{frame}{ }
  \begin{itemize}
    \item 
  \end{itemize}
\end{frame}


\end{document}
